MASU Paper

### 2022-05-27 paper/paper_draft.Rmd
updated discussion section

# Discussion 
## Results Discussion
According to the PCA, there is overlap between the anadromous and resident individuals from both populations. In other words, there are not clear distinct clusters that separate the anadromous individuals from the resident individuals. This indicates there is not strong evidence for a genetic influence that determines the MASU salmon's life cycle.

## Ramifications for MASU salmon
In these populations it is possible that residency and anadromy within MASU salmon are not solely genetically controlled and that we must take into consideration the environmental influences. When comparing MASU salmon to steelhead, steelhead and their life cycles are more heavily influenced by genetic factors. As of now, MASU salmon and their life cycles are mainly influenced my environmental factors, but further studies with larger sample sizes would need to be conducted in order to see if any genetic influences play a role. In other words, we would need still need to figure out if MASU salmon life cycles are influenced by the additive effects of genetics and environment, the multiplying effetcs of genetics and environment, or simply a result of purely environmental effects.

## Ramifications for Salmonids in General 
In terms of the importance of this work, figuring out if MASU salmon life cycles have genetics influences will go towards their conservation efforts. For example, within steelhead fish, anadromous individuals will have to migrate upstream in order to move towards the ocean. On some occurrences the anadromous steelhead will be stopped midway towards their journey to the ocean due to manmade dams. This has caused a reduction in diversity within these fish by causing an apparent increase in resident individuals within the population. Knowing whether a fish is anadromous or resident would be helpful because people can advocate for dams to built in areas where there are a lot of anadromous fish. This same logic can also be applied towards the conservation of MASU salmon.

>>>>

# Discussion 
## Results Discussion
According to the PCA, there is overlap between the anadromous and resident individuals from both populations. Although having overlapping clusters does not exclude the possibility of a genetic influence, our findings did not clearly show associated genetic elements between resident and anadromous fish. As mentioned in the previously to look for patterns in relatedness between resident and andromous individuals, we calculated the mean PC1 and PC2 coordinates for each stream, then calculated the distance between each individual sample and the mean for the other population. It is important to note that we did find that anadromous individuals were more closely related to individuals in the other population. 
<!-- This can potentially be attributed to where the samples were collected. For example, if the anadromous individuals from one stream were collected in an area that was geographically near the individuals in the other stream, this could explain this observation. (not sure how to interpret the metadata to confirm this) -->
We also made a manhattan plot and did not find many individuals that had a locus with a high p-value. This indicates that there is not an area in the genome that is strongly associated with the given trait (anadromy vs. resident). The random forest also did not have strong predictive power for anadromy and residency because it could not accurately group the individuals to their actual life cycle. 
Overall there is not strong evidence for a genetic influence that determines the MASU salmon's life cycle.

## Ramifications for MASU salmon
In these populations it is possible that residency and anadromy within MASU salmon are not solely genetica lly controlled and that we must take into consideration the environmental influences. When comparing MASU salmon to steelhead, steelhead and their life cycles are more heavily influenced by genetic factors. As of now, MASU salmon and their life cycles are mainly influenced my environmental factors, but further studies with larger sample sizes would need to be conducted in order to see if any genetic influences play a role. In other words, we would need still need to figure out if MASU salmon life cycles are influenced by the additive effects of genetics and environment, the multiplying effetcs of genetics and environment, or simply a result of purely environmental effects.

<!-- talk about genetic mechanism in steelhead and how that varies in environmental  -->
## Ramifications for Salmonids in General 
In terms of the importance of this work, figuring out if MASU salmon life cycles have genetics influences will go towards their conservation efforts. For example, within steelhead fish, anadromous individuals will have to migrate upstream in order to move towards the ocean. On some occurrences the anadromous steelhead will be stopped midway towards their journey to the ocean due to manmade dams. This has caused a reduction in diversity within these fish by causing an apparent increase in resident individuals within the population. Knowing whether a fish is anadromous or resident would be helpful because people can advocate for dams to built in areas where there are a lot of anadromous fish. This same logic can also be applied towards the conservation of MASU salmon.
